Exploring how changes in age at death and cause of death have narrowed the black-white life expectancy gap over forty-five years in the United States.

### 2017-01-05 Code/append_smoothed_results.Rmd

```{r compute_marginal_COD_decomposition}
list.cod.marginal.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") { 
       list.cod.marginal.tables.smoothed[[i]] <-  list.cod.decomp.tables.smoothed[[i]] %>%
                                                  dplyr::group_by(Cause.of.death) %>% 
                                                  dplyr::summarise(C_x_COD = sum(C_xi)) %>%
                                                  dplyr::mutate(C_x_COD_proportion = C_x_COD/sum(C_x_COD),
                                                         adds_to_gap = ifelse(C_x_COD > 0, T, F))

    }
}
```

>>>>

```{r compute_marginal_COD_decomposition}
list.cod.marginal.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])
State2 <- NULL
Year3 <- NULL
Sex2 <- NULL

for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") { 
       # list.cod.marginal.tables.smoothed[[i]] <-  list.cod.decomp.tables.smoothed[[i]] %>%
       #                                            dplyr::group_by(Cause.of.death) %>%
       #                                            dplyr::summarise(C_x_COD = sum(C_xi)) %>%
       #                                            dplyr::mutate(C_x_COD_proportion = C_x_COD/sum(C_x_COD),
       #                                                   adds_to_gap = ifelse(C_x_COD > 0, T, F))

       Sex2 <- c(Sex2, rep(as.character(paired.ids$Sex2[i]), 6))
       State2 <- c(State2, rep(as.character(paired.ids$State2[i]), 6))
       Year3 <- c(Year3, rep(paired.ids$Year3[i], 6))
    }
}
```

```{r create_dataset_COD_contributions}
cod.contributions <- do.call(rbind, list.cod.marginal.tables.smoothed)
cod.contributions$Year3 <- Year3
cod.contributions$State2 <- State2
cod.contributions$Sex2 <- Sex2
rm(Year3, State2, Sex2)

cod.contributions <- cod.contributions %>% mutate(stabbrs = factor(State2,  
                                      labels = c("AL", "AZ", "AR", "CA", "CO", "CT", "DE", 
                                                 "FL", "GA", "IL", "IN", "IA", "KS", "KY", 
                                                 "LA", "MD", "MA", "MI", "MN", "MS", "MO", 
                                                 "NE", "NV", "NJ", "NM", "NY", "NC", "OH", 
                                                 "OK", "PA", "RI", "SC", "TN", "TX", "UT", 
                                                 "VA", "WA", "DC", "WV", "WI")))
table(cod.contributions$State2, cod.contributions$stabbrs)

cod.contributions$Census_Region <- "South"
cod.contributions$Census_Region[cod.contributions$State2 %in% c("Iowa", "Kansas", "Minnesota", "Missouri", 
                                                 "Nebraska", "North Dakota", "South Dakota", 
                                                 "Illinois", "Indiana", "Michigan", "Ohio", 
                                                 "Wisconsin")] <- "Midwest"

cod.contributions$Census_Region[cod.contributions$State2 %in% c("Connecticut", "Maine", "Massachusetts", 
                                                 "New Hampshire", "Rhode Island", "Vermont",
                                                 "New Jersey", "New York", "Pennsylvania")] <- "Northeast"

cod.contributions$Census_Region[cod.contributions$State2 %in% c("Arizona", "Colorado", "Idaho", "Montana", 
                                                 "Nevada", "New Mexico", "Utah", "Wyoming",
                                                 "Alaska", "California", "Hawaii", "Oregon",
                                                 "Washington")] <- "West"

### 2017-01-05 Manuscript/Manuscript.Rmd

```{r figure5, fig.align="center", fig.height=8, fig.width=8, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Male"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap(~Census_Region) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") #+
           #geom_text(data = subset(BlackWhite, Year3 == 2010), aes(label = State2), check_overlap = T)
         )
```

@@

```{r figure6, fig.align="center", fig.height=8, fig.width=8, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Female"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap(~Census_Region))
```

**Figures 7 and 8 to be added: Trends in the contribution of CVD to the black-white life expectancy gap (in years) by US state**
-make one for males and one for females
-separately for each COD group
-included CIs

>>>>

```{r figure5, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Male"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap( ~ Census_Region) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite, Year3 == 2013 & Sex2 == "Male"), aes(label = stabbrs), check_overlap = T, size = 2.5) 
         ) 
```

@@

```{r figure6, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Female"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap(~Census_Region) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite, Year3 == 2013 & Sex2 == "Female"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7a: Contribution of cancer to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7a, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cancers"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)

#something is wrong with the percent contribution calculated for RI around 2008 ... hmmph
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD_proportion)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cancers"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7b: Contribution of cardiovascular disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7b, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)

#need to figure out what is weird about this one in NM in 2002 -- something is very odd. 
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD_proportion)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7c: Contribution of communicable disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7c, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Communicable"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7d: Contribution of non-communicable disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7d, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Non-communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Non-communicable"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7e: Contribution of injuries to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7e, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Injuries"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Injuries"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7f: Contribution of all other causes to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{figure7f, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "All other causes"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "All other causes"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8a: Contribution of cancer to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{figure8a, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Cancers"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8b: Contribution of cardiovascular disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{figure8b, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Cardiovascular"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8c: Contribution of communicable disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{figure8c, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Communicable"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8d: Contribution of non-communicable disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{figure8d, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Non-communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Non-Communicable"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

-add CIs when I have them
